Trending assessment for nowcasts, measurements, and forecasts

### 2024-04-02 paper.tex
Add mail

\title{Trending assessment for nowcasts, measurements, and forecasts}
%\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
\author*[1]{Jonas Rieger}
\author[1]{Bolin Liu}
\author[1]{Oliver Grothe}

@@

\abstract{
An intuitive and easily interpretable information on a quantity measured over time is whether it increases or decreases.
If the quantity of interest is not yet accessible, we refer to the ability to accurately predict, nowcast, or measure the direction of change for a true variable as trending. 
An appropriate assessment of the trending ability is crucial for decision-makers, as it enables them to assess the method's credibility and changes in the underlying process early.
While the current practice of evaluating signals is focused on averaging over local deviation measures, we present trending evaluation techniques as accompanying assessments that consider whether the development direction fits the quantity of interest. 
We formalize the concept of trending for all application areas and present quantitative measures. 
Using bootstrap confidence intervals and exclusion areas, we provide extensions for noisy data and estimation uncertainty.
We review and present new visualization techniques to assess the local trending behavior and enlarge the scope to probabilistic nowcasts and forecasts. 
We apply the methods to three large medical data applications: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
}

>>>>

\title{Trending assessment for nowcasts, measurements, and forecasts}
%\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
\author*[1]{Jonas Rieger}\email{jonas.rieger@kit.edu}
\author[1]{Bolin Liu}
\author[1]{Oliver Grothe}

@@

\abstract{
An intuitive and easily interpretable information on a quantity measured over time is whether it increases or decreases.
We refer to the ability to accurately predict, nowcast, or measure the direction of change of a true variable as trending. 
An appropriate assessment of the trending ability is crucial for decision-makers, as it enables them to assess the method's credibility and changes in the underlying process early.
While the current practice of evaluating signals is focused on averaging over local deviation measures, we present trending evaluation techniques as accompanying assessments that consider whether the development direction fits the quantity of interest. 
We formalize the concept of trending for all application areas and present quantitative measures. 
Using bootstrap confidence intervals and exclusion areas, we provide extensions for noisy data and estimation uncertainty.
We review and present new visualization techniques to assess the local trending behavior and enlarge the scope to probabilistic nowcasts and forecasts. 
We apply the methods to three large medical data applications: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
}
